Eyeballing Parameters
=============

For each graph, choose an appropriate function form to represent the pattern shown by the data and estimate, by eye, the parameters that will fit the data.

```{r include=FALSE}
require(mosaic)
opts_chunk$set(tidy=FALSE, fig.width=3, fig.height=3)
```


The fraction of US households with a telephone, as a function of family income.  (Source: Susan E. Mayer (1997) **What money can't buy: Family income and children's life chances** Harvard Univ. Press p. 102.)
```{r message=FALSE,fig.keep="all"}
inc = fetchData("Income-Housing.csv")
plotPoints( Telephone ~ IncomePercentile, data=inc)
```

Stan Wagon's measurements of the temperature of a water.  The water was boiled and poured into the mug.  (Source: Stan Wagon and Robert Portmann, "How quickly does hot water cool?", *Mathematica in Education and Research* **10**:3 (July 2005) 1-9.)
```{r message=FALSE}
stan = fetchData("stan-data.csv")
plotPoints( temp ~ time, data=stan)
```

Kepler's measurements of the position of Mars relative to the sun.
(Data source: McLaughlin, Michael P. ( 1999 ) ``A Tutorial on
  Mathematical Modelling'' <http://www.causascientia.org/math_stat/Tutorial.pdf> p. 21-23.)  The `Time` variable is measured as a Julian date; 0 is Greenwich noon, on 1 January 4713 BC.  Radii are in astronomical units (AU) --- the mean distance from the Earth to the Sun. 

```{r message=FALSE}
kdata = fetchData("kepler-mars.csv")
plotPoints( kepler.radius ~ kepler.angle, 
    data=kdata, 
    xlab="Angle",ylab="Radius" )
kdata = transform(kdata, x=kepler.radius*cos(kepler.angle))
kdata = transform(kdata, y=kepler.radius*sin(kepler.angle))
plotPoints(y~x, data=kdata,col="red")
```

Tidal measurements in Pearl Harbor, Hawaii.  (Source: Andrew Beveridge)

```{r message=FALSE}
h = fetchData("hawaii.csv")
plotPoints(water ~ time, data=h)
```

Data about the size, power, and other characteristics of internal combustion engines.  (Source: Thomas McMahon and John Tyler Bonner (1983) *On Size and Life*)
```{r message=FALSE}
engines = fetchData("engines.csv")
plotPoints(log(RPM)~log(BHP), data=engines)
```


Residual life expectancy versus age in the US.  US Census Bureau.  
```{r message=FALSE}
LE = fetchData("LifeExpectancy.csv")
plotPoints(LifeExpectancy ~ Age, data=LE)
```

Keeling's record of atmospheric carbon dioxide from Mauna Loa, 1958-2008. <http://cdiac.ornl.gov/trends/co2/sio-mlo.html>

```{r message=FALSE}
CO2 = fetchData("maunaloa-CO2.csv")
plotPoints( Average ~ Year, data=CO2)
```

The number of O-ring failures in each Space Shuttle launch up to the 1986 Challenger accident.

```{r message=FALSE}
oring = fetchData("oring-damage.csv")
plotPoints(damage ~ temp, data=oring, 
           pch=20, col=rgb(0,0,0,.3))
```

Data from a problem in Stewart, giving world population versus year.(Source: Stewart, **Calculus: Concepts and Contexts** 2/e p. 38.

```{r message=FALSE}
worldPop = fetchData("PREP-Stewart-World-Population.csv")
plotPoints(Population ~ Year, data=worldPop )
```


Month-by-month utility bill data from a single-family house in Saint Paul, Minnesota.
```{r message=FALSE}
utils = fetchData("utilities.csv")
plotPoints(ccf ~ temp, data=utils)
```

The radius of the fireball versus time measured at the first atomic bomb test, [Trinity](http://en.wikipedia.org/wiki/Trinity_(nuclear_test\)) in Alamogordo, New Mexico, on July 16, 1945.  
```{r message=FALSE}
blast = fetchData("blastdata.csv")
plotPoints(radius ~ time, data=blast)
```

Body circumference measurements for
252 men, along with estimates of the percentage of body fat
determined by underwater weighing. (Source: [StatLib](http://lib.stat.cmu.edu/datasets/bodyfat).  See also a discussion by [Victor Addona](http://www.mosaic-web.org/go/MCAST/materials/Sept-10-2010/Addona_MCAST_Sept_10.pdf).) 
```{r message=FALSE}
body = fetchData("BodyFat.csv")
body = subset(body, Height>60 & Weight<300 )
plotPoints(Weight ~ Height, data=body)
```

Data on used Honda Accords collected by a student group in Macalester's Math 155 class.  (Contact: Daniel Kaplan, Macalester College)
```{r message=FALSE}
cars = fetchData("used-hondas.csv")
plotPoints(Price ~ Age, data=cars)
```