Eyeballing Parameters
=============

   * Exponentials
       * Income-housing for exponential. [exercise](http://rpubs.com/dtkaplan/558)
       * `stan-data.csv` for exponential. [Is cooling exponential?](http://rpubs.com/dtkaplan/558)
    * Trigonometric
       * tides for sine wave [from PREP2012](http://prep2012.mosaic-web.org/Topics/Tides/tides-hawaii.html)
    * Power Law
       * atomic bomb radius versus time
    * S-curves
       * gaussian and its cumulative as an S-curve
       * logistic function
       * [Hill function](http://prep2012.mosaic-web.org/Topics/Functions/hillFunctions.html)
    * Gaussian bump 
        * Ben Baumer's shortstop example
    * Quadratic
        * [Kepler's orbit](Notes/Kepler-Mars.html)
        

```{r include=FALSE}
require(mosaic)
opts_chunk$set(tidy=FALSE, fig.width=3, fig.height=3)
```


```{r message=FALSE,fig.keep="all"}
inc = fetchData("Income-Housing.csv")
plotPoints( Telephone ~ IncomePercentile, data=inc)
```

```{r message=FALSE}
stan = fetchData("stan-data.csv")
plotPoints( temp ~ time, data=stan)
```

```{r message=FALSE}
kdata = fetchData("kepler-mars.csv")
plotPoints( kepler.radius ~ kepler.angle, 
    data=kdata, 
    xlab="Angle",ylab="Radius" )
kdata = transform(kdata, x=kepler.radius*cos(kepler.angle))
kdata = transform(kdata, y=kepler.radius*sin(kepler.angle))
plotPoints(y~x, data=kdata,col="red")
```

```{r message=FALSE}
h = fetchData("hawaii.csv")
plotPoints(water ~ time, data=h)
```

```{r message=FALSE}
engines = fetchData("engines.csv")
plotPoints(log(RPM)~log(BHP), data=engines)
```

```{r message=FALSE}
LE = fetchData("LifeExpectancy.csv")
plotPoints(LifeExpectancy ~ Age, data=LE)
```

```{r message=FALSE}
CO2 = fetchData("maunaloa-CO2.csv")
plotPoints( Average ~ Year, data=CO2)
```

```{r message=FALSE}
oring = fetchData("oring-damage.csv")
plotPoints(damage ~ temp, data=oring, 
           pch=20, col=rgb(0,0,0,.3))
```


```{r message=FALSE}
worldPop = fetchData("PREP-Stewart-World-Population.csv")
plotPoints(Population ~ Year, data=worldPop )
```

```{r message=FALSE}
utils = fetchData("utilities.csv")
plotPoints(ccf ~ temp, data=utils)
```

```{r message=FALSE}
blast = fetchData("blastdata.csv")
plotPoints(radius ~ time, data=blast)
```

```{r message=FALSE}
body = fetchData("BodyFat.csv")
body = subset(body, Height>60 & Weight<300 )
plotPoints(Weight ~ Height, data=body)
```

```{r message=FALSE}
cars = fetchData("used-hondas.csv")
plotPoints(Price ~ Age, data=cars)
```