Estimating Sampling Variation
==========================

```{r include=FALSE}
require(mosaic)
require(xtable)
opts_chunk$set(tidy=FALSE, fig.width=3, fig.height=3)
options(xtable.type="html")
```

### Danny Kaplan, `r I(date())`

Sampling variation refers to the uncertainty created by the sampling process.  If your data is just a sample, another data set will produce a somewhat different result.  The question we seek to answer is, "How much is *somewhat*?"

The method used here is called **bootstrapping**.  Bootstrapping relies on a simulation of the sampling process called **resampling**: sampling from the sample.

#### Start with your data and a model

To illustrate, consider estimating the power-law exponent on the relationship between species number and land area.

```{r  fig.width=3,fig.height=3,message=FALSE}
species = fetchData("http://www.math.smith.edu/~bbaumer/mth247/SpeciesArea.csv")
plotPoints(Species ~ Area, data=species)
```

The model we seek to fit is Species = a Area^b.  Judging from the shape of the graph, Species goes roughly as the square-root of area, so use $b=0.5$ as the initial guess for the nonlinear parameter.  Finding the parameters is a straightforward application of `fitModel`:

```{r}
fitModel( Species ~ a*Area^b, 
          data=species, 
          start=list(b=.5) )
```

It looks like a fourth-root model is appropriate.  But how precise is the estimate of the coefficient $b$?

#### Package up the parameter estimation

Write up the estimation calculation in a form that returns the values of interest.  Since $b$ is the parameter of interest here, we already have that form.  

But to illustrate the process, suppose that we were interested in predicting the number of species in a land with area $2 \times 10^6$ km^2.  This would be an extrapolation, and extrapolations are generally unreliable, but here goes ...

```{r}
{ mod = fitModel( Species ~ a*Area^b, 
                  data=species, 
                  start=list(b=.5) )
  mod( 2e6 ) 
}
```

Note that the pair of commands has been put between curly braces.  That's the R syntax for grouping multiple statements into a single expression.

#### Apply resampling and iterate

Take your statement, appropriately packaged.  Wrap the data frame with `resample`.  Put `do(100)*` in front of the statement and assign the results so that you can access them later:

```{r}
s = do(100)*
  { mod = fitModel( Species ~ a*Area^b, 
                  data=resample(species), 
                  start=list(b=.5) )
              coef(mod) 
  }
```
Now you have a set of 100 resampling realizations.  This can be handled in stand ways.  For instance, to find the 95% confidence interval, use `confint`.

```{r}
confint(s)
```

Returning to the coefficient $b$ ...
```{r}
s2 = do(100)*
   {mod = fitModel( Species ~ a*Area^b, 
                  data=resample(species), 
                  start=list(b=.5) )
    coef(mod)}
```
And the confidence intervals:
```{r}
confint(s2)
```